TODO LIST

### 2019-01-21 LMCS/sendrev/syntactic-clo.tex
reviewer 1 - technical casi listo
varias reescrituras gramaticales

We use the \emph{$\eqso(\fo)$-hierarchy} to define syntactic classes of functions with good algorithmic and closure properties.  
But before doing this, we introduce a more strict notion of counting problem with easy decision version.
Recall that a function $f : \Sigma^* \to \mathbb{N}$ has an easy decision counterpart if $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$ is a language in $\ptime$. As the goal of this section is to define a syntactic class of functions in $\shp$ with easy decision versions and good closure properties, we do not directly consider the semantic condition $L_f  \in \ptime$, but instead we consider a more restricted syntactic condition. More precisely, a function $f : \Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$~\cite{PagourtzisZ06} if there exists a  polynomial-time NTM $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$. Notice that one is subtracted from $\tmt_M(x)$ to allow for $f(x) = 0$. Besides, notice that $\totp \subseteq \shp$ and that $f \in \totp$ implies that $L_f \in \ptime$.

@@

It was proved in \cite{SalujaST95} that every function in $\E{1}$ admits an FPRAS. Besides, it can be proved that $\E{1} \subseteq \totp$. 
However, this class is not closed under sum, and then it is not robust under basic closure properties. 
\begin{prop}\label{prop-e1-nc}
There exist functions $f, g \in \E{1}$ such that $(f + g) \not\in \E{1}$.
\end{prop}
\proof
\input{proof-prop-e1-nc}
\qed
To overcome this limitation, one can consider the class $\QE{1}$, which is closed under sum by definition. In fact, the following proposition shows that the same good properties as for $\E{1}$ hold for $\QE{1}$, together with the fact that it is closed under sum and multiplication.
\begin{prop} \label{prop:qe0-fp-qe1-totp-fptras}
$\QE{1} \subseteq \totp$ and every function in $\QE{1}$ has an FPRAS. Moreover, $\QE{1}$ is closed under sum and multiplication.
\end{prop}
\proof
\input{proof-qe0-fp-qe1-totp-fptras}
\qed
Hence, it only remains to prove that $\QE{1}$ is closed under subtraction by one. Unfortunately, it is not clear whether this property holds; in fact, we conjecture that it is not the case. Thus, we need to find an extension of $\QE{1}$ that keeps all the previous properties and is closed under subtraction by one. It is important to notice that $\shp$ is believed not to be closed under subtraction by one due to some complexity-theoretical assumption\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a polynomial-time NTM $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}. It is believed that $\np \not\subseteq \cspp$.
However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.}. So, the following proposition rules out any logic that extends $\Pi_1$ as a possible extension of~$\QE{1}$ with the desired closure property.
\begin{prop} \label{pi-minusone}
If $\Pi_1 \subseteq \LL \subseteq \fo$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by~one. 
\end{prop}
\proof
\input{proof-pi-minusone}
\qed
Therefore, the desired extension has to be achieved by allowing some local extensions to~$\Sigma_1$. More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature~$\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mention any second-order variable). With this extension we obtain a class with the desired properties.
\begin{thm}\label{sigmafo-minusone}
The class $\eqso(\logex{1})$ is closed under sum, multiplication and subtraction by one. Moreover, $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
\end{thm}
\proof
\input{proof-sigmafo-minusone}
\qed
The proof that $\eqso(\logex{1})$ is closed under subtraction by one is the most involved of the paper. We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.

>>>>

We use the \emph{$\eqso(\fo)$-hierarchy} to define syntactic classes of functions with good algorithmic and closure properties.  
But before doing this, we introduce a more strict notion of counting problem with easy decision version.
Recall that a function $f\colon\Sigma^* \to \mathbb{N}$ has an easy decision counterpart if $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$ is a language in $\ptime$. As the goal of this section is to define a syntactic class of functions in $\shp$ with easy decision versions and good closure properties, we do not directly consider the semantic condition $L_f  \in \ptime$, but instead we consider a more restricted syntactic condition. More precisely, a function $f\colon\Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$~\cite{PagourtzisZ06} if there exists a  polynomial-time NTM $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$. Notice that one is subtracted from $\tmt_M(x)$ to allow for $f(x) = 0$. Besides, notice that $\totp \subseteq \shp$ and that $f \in \totp$ implies that $L_f \in \ptime$.

@@

It was proved in \cite{SalujaST95} that every function in $\E{1}$ admits an FPRAS. Besides, it can be proved that $\E{1} \subseteq \totp$. 
However, this class is not closed under sum, so it is not robust under the basic closure properties we are looking for. 
\begin{prop}\label{prop-e1-nc}
There exist functions $f, g \in \E{1}$ such that $(f + g) \not\in \E{1}$.
\end{prop}
\proof
\input{proof-prop-e1-nc}
\qed
To overcome this limitation, one can consider the class $\QE{1}$, which is closed under sum by definition. In fact, the following proposition shows that the same good properties as for $\E{1}$ hold for $\QE{1}$, together with the fact that it is closed under sum and multiplication.
\begin{prop} \label{prop:qe0-fp-qe1-totp-fptras}
$\QE{1} \subseteq \totp$ and every function in $\QE{1}$ has an FPRAS. Moreover, $\QE{1}$ is closed under sum and multiplication.
\end{prop}
\proof
\input{proof-qe0-fp-qe1-totp-fptras}
\qed
Hence, it only remains to prove that $\QE{1}$ is closed under subtraction by one. Unfortunately, it is not clear whether this property holds; in fact, we conjecture that it is not the case. Thus, we need to find an extension of $\QE{1}$ that keeps all the previous properties and is closed under subtraction by one. It is important to notice that $\shp$ is believed not to be closed under subtraction by one due to some complexity-theoretical assumption\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a polynomial-time NTM $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}. It is believed that $\np \not\subseteq \cspp$.
However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.}. So, the following proposition rules out any logic that extends $\Pi_1$ as a possible extension of~$\QE{1}$ with the desired closure property.
\begin{prop} \label{pi-minusone}
If $\Pi_1 \subseteq \LL \subseteq \fo$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by~one. 
\end{prop}
\proof
\input{proof-pi-minusone}
\qed
Therefore, the desired extension has to be achieved by allowing some local extensions to~$\Sigma_1$. To this end, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature~$\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mention any second-order variable). With this extension we obtain a class with the desired properties.
\begin{thm}\label{sigmafo-minusone}
The class $\eqso(\logex{1})$ is closed under sum, multiplication and subtraction by one. Moreover, $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
\end{thm}
\begin{leftbar}
\proof
\input{proof-sigmafo-minusone}
\qed
\end{leftbar}
The proof that $\eqso(\logex{1})$ is closed under subtraction by one is the most involved of the paper. We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.
\martin{Esta correcci\'on se me fue en collera}

### 2019-01-21 LMCS/sendrev/syntactic.tex
varias reescrituras gramaticales

The importance of the class of counting problems with easy decision counterparts has motivated the search for robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}. But the key question here is what should be considered a {\em robust} class. 
A first desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}. Analogously to the cases of $\ptime$ and $\np$, which are closed under intersection and union, we expect our class to be closed under multiplication and sum. For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula. Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}}$ is trivial. But should $\textit{sat\_one}$ be part of a robust class of counting functions with easy decision versions? The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for~$x$. 
%As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula. 
Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them. In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g \dotdiv 1$ should also be in $\CC$, where $(g \dotdiv 1)(x)$ is defined as $g(x) - 1$ if $g(x) \geq 1$, and as~$0$ otherwise. Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} \dotdiv 1$ counts the number of satisfying assignments of a propositional formula.

>>>>

The importance of the class of counting problems with easy decision counterparts has motivated the search for robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}. But the key question here is what should be considered a {\em robust} class. 
A first desirable condition is related to its closure properties, which is a common theme when studying function complexity classes \cite{OH93,FH08}. Analogously to the cases of $\ptime$ and $\np$, which are closed under intersection and union, we expect our class to be closed under multiplication and sum. For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula. Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}}$ is trivial. But should $\textit{sat\_one}$ be part of a robust class of counting functions with easy decision versions? The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for~$x$. 
%As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula. 
Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them. In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g \dotdiv 1$ should also be in $\CC$, where $(g \dotdiv 1)(x)$ is defined as $g(x) - 1$ if $g(x) \geq 1$, and as~$0$ otherwise. Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} \dotdiv 1$ counts the number of satisfying assignments of a propositional formula.

### 2019-01-21 LMCS/sendrev/todo-list.tex
varias reescrituras gramaticales

	\item generally: both formulae and formulas are used as plural of formula---should be consistent
	\item p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict
	WL to the natural numbers as fixed semiring, calling this restriction ...''

@@

	\item p5, description of syntax of QSO: ``This division between Boolean and quantitative level
	is essential for understanding the difference between the logic and the quantitative part.''
	reformulate this?

@@

	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence

@@

	\item throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$

@@

	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is

@@

	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise

@@

	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''

@@

	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'':

@@

	\item p.19: ``However, this class is not closed under sum, and then it is not robust under basic

@@

	\item p.21: ``the desired extension has to be achieved by ... . More precisely, we define...'': The second

@@

	\item p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated

@@

	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there

@@

	\item on definition of functions, \verb|\colon| should be used instead of :

@@

\item[$\checkmark$] page 4:
-some of the $\to$ some of them 
\item[$\checkmark$] page 10: 
-$2^{|A|^l}$-bits $\to$ $2^{|A|^l}$ bits
\item[$\checkmark$] page 21:
-it does not mentioned $\to$ it does not mention
\item page 22: 
-you make a long proof about some closure under subtraction by one.
You should put the result in a theorem envirpnment to make what you are
proving more evident.
\item[$\checkmark$] -replace .... for $\to$ replace ... by;  there are several instances of
that
\item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
\item -you explain that you rewrite your formula to obtain conditions a,b and c, 
you should give an idea how these conditions will be used
\item[$\checkmark$] page 23:
a SO, a FO  an SO, an FO
\item page 24: 
if, if $\phi$$\to$ $\phi$
\item page 25: 
-you explain that the proof relies on witnesses of logarithmic size.
It was not clear to me while reading the proof. Can you explain more ?
\item page 28:
-the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn. You should mention here that you do not need sum here
and that the framework of [39] is enough for this result.
\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''

>>>>

	\item[$\checkmark$] generally: both formulae and formulas are used as plural of formula---should be consistent
	\item[$\checkmark$] p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict WL to the natural numbers as fixed semiring, calling this restriction ...''

@@

	\item[$\checkmark$] p5, description of syntax of QSO: ``This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part.'' reformulate this?

@@

	\item[$\checkmark$] p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence

@@

	\item[$\checkmark$] throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$

@@

	\item[$\checkmark$] p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is

@@

	\item[$\checkmark$] p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise

@@

	\item[$\checkmark$] p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''

@@

	\item[$\checkmark$] p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'':

@@

	\item[$\checkmark$] p.19: ``However, this class is not closed under sum, and then it is not robust under basic

@@

	\item[$\checkmark$] p.21: ``the desired extension has to be achieved by ... . More precisely, we define...'': The second

@@

	\item[$\checkmark$] p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated

@@

	\item[$\checkmark$] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there

@@

	\item[$\checkmark$] on definition of functions, \verb|\colon| should be used instead of :

@@

\item[$\checkmark$] page 4:
-some of the $\to$ some of them 
\item[$\checkmark$] page 10: 
-$2^{|A|^l}$-bits $\to$ $2^{|A|^l}$ bits
\item[$\checkmark$] page 21:
-it does not mentioned $\to$ it does not mention
\item[$\checkmark$] page 22: 
-you make a long proof about some closure under subtraction by one.
You should put the result in a theorem envirpnment to make what you are
proving more evident.
\item[$\checkmark$] -replace .... for $\to$ replace ... by;  there are several instances of
that
\item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
\item -you explain that you rewrite your formula to obtain conditions a,b and c, 
you should give an idea how these conditions will be used
\item[$\checkmark$] page 23:
a SO, a FO  an SO, an FO
\item page 24: 
if, if $\phi$$\to$ $\phi$
\item page 25: 
-you explain that the proof relies on witnesses of logarithmic size.
It was not clear to me while reading the proof. Can you explain more ?
\item page 28:
-the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn. You should mention here that you do not need sum here
and that the framework of [39] is enough for this result.
\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
